Tropical Andes Taxonomic Diversity of Plants and Frugivores

### 2024-05-15 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
in richness plotting function: add direction for palettes and remove title; update exponential model equation, create seperate scatterplots for exponential and linear models, add return list of plots that accidentially was removed

  # Helper functions to generate plots
  generate_plant_plot <- function(data) {
    ggplot(data) +
      geom_sf(data = Americas, fill = "white") +
      geom_sf(data = TApoly, fill = "grey", size = 0.1) +
      geom_sf(aes(fill = num_species), color = NA) +
      scale_fill_distiller(palette = "Greens", limits = c(0,6000)) +
      labs(title = title, x = "Longitude", y = "Latitude", fill = "Number of species") +
      coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
      annotation_scale(location = "bl", width_hint = 0.5)  +
      theme(panel.background = element_rect(fill = "lightblue"))
  }
  generate_frugivore_plot <- function(data) {
    ggplot(data) +
      geom_sf(data = Americas, fill = "white") +
      geom_sf(data = TApoly, fill = "grey", size = 0.1) +
      geom_sf(aes(fill = num_species), color = NA) +
      scale_fill_distiller(palette = "YlOrBr", limits = c(0,500)) + 
      labs(title = title, x = "Longitude", y = "Latitude", fill = "Number of species") +
      coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
      annotation_scale(location = "bl", width_hint = 0.5)  +
      theme(panel.background = element_rect(fill = "lightblue"))
  }

@@

  ## Fit an exponential model
  compare_richness_exponential_summary <- summary(lm(exp(comparison_data_filtered$Frugivore_Richness) ~
                                                       comparison_data_filtered$Plant_Richness))
  exponential_r_squared <- compare_richness_exponential_summary$r.squared

  ## Create scatter plot with linear and exponential lines
  plant_frug_richness_compare <- ggplot(comparison_data,
                                        aes(x = Plant_Richness, y = Frugivore_Richness)) +
  geom_point() +
  labs(title = paste("Plant Richness vs Frugivore Richness [", resolution_meters/1000, " km]", sep = ""),
       subtitle =  paste("Linear (Red) R-squared = ", round(linear_r_squared, 2), "\n",
                  "Exponential (Blue) R-squared = ", round(exponential_r_squared, 2)),
       x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
    geom_smooth(method = "lm", se = FALSE, color = "red") +
    geom_smooth(method = "lm", se = FALSE, color = "blue", formula = y ~ exp(x), linetype = "dashed")

  # Create histograms
  plant_richness_hist <- hist(comparison_data$Plant_Richness, 
    main = paste("Histogram of Plant Richness [", resolution_meters/1000, " km] by cell", sep = ""),
                              xlab = "Plant Richness by cell")
  frugivore_richness_hist <- hist(comparison_data$Frugivore_Richness,
    main = paste("Histogram of Frugivore Richness [", resolution_meters/1000, " km] by cell", sep = ""),
                                  xlab = "Frugivore Richness by cell")
}

@@

# save plots
plantgridRichTA_1km <- richness_1km$plant_plot
frugivoregridRichTA_1km <- richness_1km$frugivoregridRichTA
# save richness data
plant_cellRichness_1km <- richness_1km$plant_cellRichness
frugivore_cellRichness_1km <- richness_1km$frugivore_cellRichness

compare_richness_r2_1km <- richness_1km$compare_richness_summary$r.squared
```

@@

# save plots
plantgridRichTA_5km <- richness_5km$plantgridRichTA
frugivoregridRichTA_5km <- richness_5km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_5km <- richness_5km$plant_cellRichness
frugivore_cellRichness_5km <- richness_5km$frugivore_cellRichness

compare_richness_r2_5km <- richness_5km$compare_richness_summary$r.squared
```

@@

# save plots
plantgridRichTA_10km <- richness_10km$plantgridRichTA
frugivoregridRichTA_10km <- richness_10km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_10km <- richness_10km$plant_cellRichness
frugivore_cellRichness_10km <- richness_10km$frugivore_cellRichness

compare_richness_r2_10km <- richness_10km$compare_richness_summary$r.squared
```

@@

# save plots
plantgridRichTA_25km <- richness_25km$plantgridRichTA
frugivoregridRichTA_25km <- richness_25km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_25km <- richness_25km$plant_cellRichness
frugivore_cellRichness_25km <- richness_25km$frugivore_cellRichness

compare_richness_r2_25km <- richness_25km$compare_richness_summary$r.squared
```

## 50 km
```{r}
richness_50km <- create_rich_plots(50000)
print(richness_50km)
# save plots
plantgridRichTA_50km <- richness_50km$plantgridRichTA
frugivoregridRichTA_50km <- richness_50km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_50km <- richness_50km$plant_cellRichness
frugivore_cellRichness_50km <- richness_50km$frugivore_cellRichness

compare_richness_r2_50km <- richness_50km$compare_richness_summary$r.squared
```

## 75 km 
```{r}
richness_75km <- create_rich_plots(75000)
print(richness_75km)
# save plots
plantgridRichTA_75km <- richness_75km$plantgridRichTA
frugivoregridRichTA_75km <- richness_75km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_75km <- richness_75km$plant_cellRichness
frugivore_cellRichness_75km <- richness_75km$frugivore_cellRichness

compare_richness_r2_75km <- richness_75km$compare_richness_summary$r.squared
```

## 100 km 
```{r}
richness_100km <- create_rich_plots(100000)
print(richness_100km)
# save plots
plantgridRichTA_100km <- richness_100km$plantgridRichTA
frugivoregridRichTA_100km <- richness_100km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_100km <- richness_100km$plant_cellRichness
frugivore_cellRichness_100km <- richness_100km$frugivore_cellRichness

compare_richness_r2_100km <- richness_100km$compare_richness_summary$r.squared
```

>>>>

  # Helper functions to generate plots
  generate_plant_plot <- function(data) {
    ggplot(data) +
      geom_sf(data = Americas, fill = "white") +
      geom_sf(data = TApoly, fill = "grey", size = 0.1) +
      geom_sf(aes(fill = num_species), color = NA) +
      scale_fill_distiller(palette = "Greens", direction = 1, limits = c(0,6000)) +
      labs( x = "Longitude", y = "Latitude", fill = "Number of species") +
      coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
      annotation_scale(location = "bl", width_hint = 0.5)  +
      theme(panel.background = element_rect(fill = "lightblue"))
  }
  generate_frugivore_plot <- function(data) {
    ggplot(data) +
      geom_sf(data = Americas, fill = "white") +
      geom_sf(data = TApoly, fill = "grey", size = 0.1) +
      geom_sf(aes(fill = num_species), color = NA) +
      scale_fill_distiller(palette = "YlOrBr", direction = 1, limits = c(0,500)) + 
      labs(x = "Longitude", y = "Latitude", fill = "Number of species") +
      coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
      annotation_scale(location = "bl", width_hint = 0.5)  +
      theme(panel.background = element_rect(fill = "lightblue"))
  }

@@

  ## Fit an exponential model
  compare_richness_exponential_summary <- summary(lm(comparison_data_filtered$Frugivore_Richness ~
                                                       log(comparison_data_filtered$Plant_Richness)))
  exponential_r_squared <- compare_richness_exponential_summary$r.squared

  ## Create scatter plot with linear and exponential lines
  compare_richness_linear_plot <- ggplot(comparison_data,
                                        aes(x = Plant_Richness, y = Frugivore_Richness)) +
  geom_point() +
  labs(title = paste("Plant Richness vs Frugivore Richness [", resolution_meters/1000, " km]", sep = ""),
       subtitle =  paste("Linear (Red) R-squared = ", round(linear_r_squared, 2)),
       x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
    geom_smooth(method = "lm", se = FALSE, color = "red")

  compare_richness_exponential_plot <- ggplot(comparison_data,
                                        aes(x = Plant_Richness, y = Frugivore_Richness)) +
  geom_point() +
  labs(title = paste("Plant Richness vs Frugivore Richness [", resolution_meters/1000, " km]", sep = ""),
       subtitle =  paste("Exponential (Blue) R-squared = ", round(exponential_r_squared, 2)),
       x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
    geom_smooth(method = "lm", se = FALSE, color = "blue", formula = y ~ log(x))

  # Create histograms
  plant_richness_hist <- hist(comparison_data$Plant_Richness, 
    main = paste("Histogram of Plant Richness [", resolution_meters/1000, " km] by cell", sep = ""),
                              xlab = "Plant Richness by cell")
  frugivore_richness_hist <- hist(comparison_data$Frugivore_Richness,
    main = paste("Histogram of Frugivore Richness [", resolution_meters/1000, " km] by cell", sep = ""),
                                  xlab = "Frugivore Richness by cell")
  # Return a list of plots
  list(plantgridRichTA = plant_plot,
       frugivoregridRichTA = frugivore_plot,
       plant_cellRichness = plant_richness_grid,
       frugivore_cellRichness = frugivore_richness_grid, 
       compare_richness_linear_summary = compare_richness_linear_summary,
       compare_richness_exponential_summary = compare_richness_exponential_summary,
       compare_richness_linear_plot = compare_richness_linear_plot,
       compare_richness_exponential_plot = compare_richness_exponential_plot,
       plant_richness_hist = plant_richness_hist,
       frugivore_richness_hist = frugivore_richness_hist)
}

@@

# save plots
plantgridRichTA_1km <- richness_1km$plant_plot
frugivoregridRichTA_1km <- richness_1km$frugivoregridRichTA
# save richness data
plant_cellRichness_1km <- richness_1km$plant_cellRichness
frugivore_cellRichness_1km <- richness_1km$frugivore_cellRichness
# save linear and exponential summaries
compare_richness_linear_summary_1km <- richness_1km$compare_richness_linear_summary
compare_richness_exponential_summary_1km <- richness_1km$compare_richness_exponential_summary
# save r-squared for linear and exponential models
compare_richness_linear_r2_1km <- richness_1km$compare_richness_linear_summary$r.squared
compare_richness_exponential_r2_1km <- richness_1km$compare_richness_exponential_summary$r.squared
```

@@

# save plots
plantgridRichTA_5km <- richness_5km$plantgridRichTA
frugivoregridRichTA_5km <- richness_5km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_5km <- richness_5km$plant_cellRichness
frugivore_cellRichness_5km <- richness_5km$frugivore_cellRichness
# save linear and exponential summaries
compare_richness_linear_summary_5km <- richness_5km$compare_richness_linear_summary
compare_richness_exponential_summary_5km <- richness_5km$compare_richness_exponential_summary
# save r-squared for linear and exponential models
compare_richness_linear_r2_5km <- richness_5km$compare_richness_linear_summary$r.squared
compare_richness_exponential_r2_5km <- richness_5km$compare_richness_exponential_summary$r.squared
```

@@

# save plots
plantgridRichTA_10km <- richness_10km$plantgridRichTA
frugivoregridRichTA_10km <- richness_10km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_10km <- richness_10km$plant_cellRichness
frugivore_cellRichness_10km <- richness_10km$frugivore_cellRichness
# save linear and exponential summaries
compare_richness_linear_summary_10km <- richness_10km$compare_richness_linear_summary
compare_richness_exponential_summary_10km <- richness_10km$compare_richness_exponential_summary
# save r-squared for linear and exponential models
compare_richness_linear_r2_10km <- richness_10km$compare_richness_linear_summary$r.squared
compare_richness_exponential_r2_10km <- richness_10km$compare_richness_exponential_summary$r.squared
```

@@

# save plots
plantgridRichTA_25km <- richness_25km$plantgridRichTA
frugivoregridRichTA_25km <- richness_25km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_25km <- richness_25km$plant_cellRichness
frugivore_cellRichness_25km <- richness_25km$frugivore_cellRichness
# save linear and exponential summaries
compare_richness_linear_summary_25km <- richness_25km$compare_richness_linear_summary
compare_richness_exponential_summary_25km <- richness_25km$compare_richness_exponential_summary
# save r-squared for linear and exponential models
compare_richness_linear_r2_25km <- richness_25km$compare_richness_linear_summary$r.squared
compare_richness_exponential_r2_25km <- richness_25km$compare_richness_exponential_summary$r.squared
```

## 50 km
```{r}
richness_50km <- create_rich_plots(50000)
print(richness_50km)
# save plots
plantgridRichTA_50km <- richness_50km$plantgridRichTA
frugivoregridRichTA_50km <- richness_50km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_50km <- richness_50km$plant_cellRichness
frugivore_cellRichness_50km <- richness_50km$frugivore_cellRichness
# save linear and exponential summaries
compare_richness_linear_summary_50km <- richness_50km$compare_richness_linear_summary
compare_richness_exponential_summary_50km <- richness_50km$compare_richness_exponential_summary
# save r-squared for linear and exponential models
compare_richness_linear_r2_50km <- richness_50km$compare_richness_linear_summary$r.squared
compare_richness_exponential_r2_50km <- richness_50km$compare_richness_exponential_summary$r.squared
```

## 75 km 
```{r}
richness_75km <- create_rich_plots(75000)
print(richness_75km)
# save plots
plantgridRichTA_75km <- richness_75km$plantgridRichTA
frugivoregridRichTA_75km <- richness_75km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_75km <- richness_75km$plant_cellRichness
frugivore_cellRichness_75km <- richness_75km$frugivore_cellRichness
# save linear and exponential summaries
compare_richness_linear_summary_75km <- richness_75km$compare_richness_linear_summary
compare_richness_exponential_summary_75km <- richness_75km$compare_richness_exponential_summary
# save r-squared for linear and exponential models
compare_richness_linear_r2_75km <- richness_75km$compare_richness_linear_summary$r.squared
compare_richness_exponential_r2_75km <- richness_75km$compare_richness_exponential_summary$r.squared
```

## 100 km 
```{r}
richness_100km <- create_rich_plots(100000)
print(richness_100km)
# save plots
plantgridRichTA_100km <- richness_100km$plantgridRichTA
frugivoregridRichTA_100km <- richness_100km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_100km <- richness_100km$plant_cellRichness
frugivore_cellRichness_100km <- richness_100km$frugivore_cellRichness
# save linear and exponential summaries
compare_richness_linear_summary_100km <- richness_100km$compare_richness_linear_summary
compare_richness_exponential_summary_100km <- richness_100km$compare_richness_exponential_summary
# save r-squared for linear and exponential models
compare_richness_linear_r2_100km <- richness_100km$compare_richness_linear_summary$r.squared
compare_richness_exponential_r2_100km <- richness_100km$compare_richness_exponential_summary$r.squared
```
